### 2025-03-04 Exam Assignments/Chapters/07/07.tex
Refining all chapters

\begin{document}
\newcommand{\drawcells}[3]{%
    \begin{tikzpicture}
        \foreach \j in {1,...,#1} {
            % Draw each cell
            \node[draw, fill=#3!20, minimum width = \linewidth/#1, minimum height=0.5cm, align=center] (cell\j) at (\j*\linewidth/#1, 0) {#2};
        }
    \end{tikzpicture}}
\section{Guided Vectorization and Data Types for Vector Intrinsics}
\subsection{Vectorization Clauses}
\begin{description}
    \item[aligned] ~\\ This clause tells the compiler that every element of an array $a$ is $n$-bit aligned if one uses \texttt{\#pragma opm simd aligned(a, n)}. The compiler employs this information to effectively vectorize the code section.
    \bigskip
    \item[safelen] ~\\ Imagine we had the following code:
    \begin{lstlisting}
a[0] = b[0];
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}
    We have a data dependency, but only every second iteration: iteration $i$, where $i$ is odd, depends on $i-1$. Iteration $i$ and $i+1$ are completely independent, however. Hence, we can vectorize the loop with a vector length of $2$. To specify this, we use the \texttt{safelen} clause like so:
    \begin{lstlisting}
a[0] = b[0];
#pragma omp simd safelen(2)
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}   
    \bigskip
    \item[reduction] ~\\ This clause has the same functionality when paired with \texttt{simd} as when paired with \texttt{paraller for}: It will reduce a variable by an associative operation. But instead of reducing it over multiple threads, we use vectorization for reducing. I imagine that the reduced variable itself will be vectorized, so that the in loop we can employ vector operations, and at the end the elements in the resulting vector will get reduced.
\end{description}

\newpage
\subsubsection{AVX512 data types}
\resizebox{\textwidth}{!}{ % Scale to the width of the text
\begin{tabular}{ r c l }
    \texttt{\_\_m512} & \drawcells{16}{F}{blue} & \texttt{16x 32-bit float} \\
    \texttt{\_\_m512d} & \drawcells{8}{D}{green} & \texttt{8x 64-bit double} \\
    \texttt{\_\_m512i} & \drawcells{47}{}{red} & \texttt{64x 8-bit byte} \\
    \texttt{\_\_m512i} & \drawcells{32}{s}{red} & \texttt{32x 16-bit short} \\
    \texttt{\_\_m512i} & \drawcells{16}{i}{red} & \texttt{16x 32-bit integer} \\
    \texttt{\_\_m512i} & \drawcells{8}{ll}{red} & \texttt{8x 64-bit long} \\
    \texttt{\_\_m512i} & \drawcells{4}{dqw}{red} & \texttt{4x 128-bit quad} \\
    \texttt{\_\_m512i} & \drawcells{2}{qqw}{red} & \texttt{2x 256-bit integer} \\
    \texttt{\_\_m512i} & \drawcells{1}{dqqw}{red} & \texttt{1x 512-bit integer} \\
\end{tabular}
}

\bigskip
\subsubsection{Intel MMX}
In the paper \href{https://dl.acm.org/doi/pdf/10.1145/242857.242865}{Intel MMX for Multimedia PCs} the authors introduced the new MMX technology, which they integrated into their new cpus. It allows the cpu to employ vector instructions to increase throuphput and hence optimize running time of common computing tasks, such as rendering pixels for example. They illustrated the use cases in several examples, two of them being:

>>>>

\begin{document}
\newcommand{\drawcells}[3]{%
    \begin{tikzpicture}
        \foreach \j in {1,...,#1} {
            % Draw each cell
            \node[draw, fill=#3!20, minimum width = \linewidth/#1, minimum height=0.5cm, align=center] (cell\j) at (\j*\linewidth/#1, 0) {#2};
        }
    \end{tikzpicture}}
\section{Guided Vectorization and Data Types for Vector Intrinsics}
\subsection{Vectorization Clauses}
\begin{description}
    \item[aligned] ~\\ This clause tells the compiler that every element of an array $a$ is $n$-byte aligned if one uses \texttt{\#pragma opm simd aligned(a, n)}. The compiler employs this information to effectively vectorize the code section.
    \bigskip
    \item[safelen] ~\\ Imagine we had the following code:
    \begin{lstlisting}
a[0] = b[0];
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}
    We have a data dependency, but only every second iteration: iteration $i$, where $i$ is odd, depends on $i-1$. Iteration $i$ and $i+1$ are completely independent, however. Hence, we can vectorize the loop with a vector length of $2$. To specify this, we use the \texttt{safelen} clause like so:
    \begin{lstlisting}
a[0] = b[0];
#pragma omp simd safelen(2)
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}   
    \bigskip
    \item[reduction] ~\\ This clause has the same functionality when paired with \texttt{simd} as when paired with \texttt{paraller for}: It will reduce a variable by an associative operation. But instead of reducing it over multiple threads, we use vectorization for reducing. I imagine that the reduced variable itself will be vectorized, so that the in loop we can employ vector operations, and at the end the elements in the resulting vector will get reduced.
\end{description}

\newpage
\subsubsection{AVX512 data types}
\begin{figure}[h]
\resizebox{\textwidth}{!}{ % Scale to the width of the text
\begin{tabular}{ r c l }
    \texttt{\_\_m512} & \drawcells{16}{F}{blue} & \texttt{16x 32-bit float} \\
    \texttt{\_\_m512d} & \drawcells{8}{D}{green} & \texttt{8x 64-bit double} \\
    \texttt{\_\_m512i} & \drawcells{47}{}{red} & \texttt{64x 8-bit byte} \\
    \texttt{\_\_m512i} & \drawcells{32}{s}{red} & \texttt{32x 16-bit short} \\
    \texttt{\_\_m512i} & \drawcells{16}{i}{red} & \texttt{16x 32-bit integer} \\
    \texttt{\_\_m512i} & \drawcells{8}{ll}{red} & \texttt{8x 64-bit long} \\
    \texttt{\_\_m512i} & \drawcells{4}{dqw}{red} & \texttt{4x 128-bit quad} \\
    \texttt{\_\_m512i} & \drawcells{2}{qqw}{red} & \texttt{2x 256-bit integer} \\
    \texttt{\_\_m512i} & \drawcells{1}{dqqw}{red} & \texttt{1x 512-bit integer} \\
\end{tabular}
}
\caption{Illustration of AVX512 data types. \texttt{\_\_m512i} is not depicted precisely.}
\end{figure}

\bigskip
\subsubsection{Intel MMX}
In the paper \href{https://dl.acm.org/doi/pdf/10.1145/242857.242865}{Intel MMX for Multimedia PCs} the authors introduced the new MMX technology, which they integrated into their new CPUs. It allows the CPU to employ vector instructions to increase throughput and hence optimize running time of common computing tasks, such as rendering pixels for example. They illustrated the use cases in several examples, two of them being:

### 2025-03-04 Exam Assignments/Chapters/09/09.tex
Refining all chapters

Now, we differentiate between two kinds of operations: Usual cpu instructions and memory operations. The reason for this is that memory operations are relatively much more time expensive, and the recent hardware trends make this gap grow bigger. Hence, we would like to know whether our program is bottlenecked by these expensive memory operations, which we very much would like to avoid, as we want to utilize our cpu with useful instructions instead.

@@

If our program is not memory-bound, it still executes not instantly, as it is bound by the execution speed of the cpu. We have a \em compute-bound\em . In a sense, every program or operation experiences a compute-bound. Hence, it is more useful when we quantify this bound, by analyzing the cpu frequency and the execution speed of individual operations, so that we can estimate the program's execution time.

@@

This is, because a lot of data will read from and written to memory, which, as explained, is very time consuming. Luckily, computer engineers introduced caches to this end, which store some data of main memory, and provide it to the program with faster access times. The only problem will be if the caches or, more precisely, their hash bucket are full. Because caches are usually not \em fully-associative\em , which means that if we have a cache line in memory, we can not write it to any cache line in the cache as we like. In fact, we only have $N$ cache lines available in an \em N-way associative \em cache.

@@

\begin{description}
    \item[Temporal locality] describes the concept that recently accessed memory has a higher probability to be accessed again.
    \item[Spatial locality] on the other hand is used to refer to the concept that memory, which is spatially close to recently accessed memory, has a higher probability of being accessed themself.
\end{description}

But why is all that useful? Well, it is due to the \em replacement policy \em of caches. As explained, caches will run full, and at that instance, they will have to evict a cache line. But how do they decide? Well, one common replacement policy is \em LRU (Least Recently Used)\em , which means that they choose the cache line whose access is the most far away in time. This is why temporal locality comes in handy, as it increases the chances that our data lies in cache, and thus reduces the risk of a cache miss.

@@

    Both matrices $a$ and $b$ are stored in row-order fashion. Hence, when running the left algorithm, $a$ has good spatial locality, whereas $b$'s is horrendous, as it leads to a cache miss on every access. The idea is now to find a tradeoff between these two. Let us analyze the access pattern of the two algorithms:

>>>>

Now, we differentiate between two kinds of operations: Usual CPU instructions and memory operations. The reason for this is that memory operations are relatively much more time expensive, and the recent hardware trends make this gap grow bigger. Hence, we would like to know whether our program is bottlenecked by these expensive memory operations, which we very much would like to avoid, as we want to utilize our CPU with useful instructions instead.

@@

If our program is not memory-bound, it still executes not instantly, as it is bound by the execution speed of the CPU. We have a \em compute-bound\em . In a sense, every program or operation experiences a compute-bound. Hence, it is more useful when we quantify this bound, by analyzing the CPU frequency and the execution speed of individual operations, so that we can estimate the program's execution time.

@@

This is, because a lot of data will be read from and written to memory, which, as explained, is very time consuming. Luckily, computer engineers introduced caches to this end, which store some data of main memory, and provide it to the program with faster access times. The only problem will be if the caches, or, more precisely, their hash bucket are full. Because caches are usually not \em fully-associative\em , which means that if we have a cache line in memory, we can not write it to any cache line in the cache as we like. In fact, we only have $N$ cache lines available in an \em N-way associative \em cache.

@@

\begin{description}
    \item[Temporal locality] describes the concept that recently accessed memory has a higher probability to be accessed again.
    \item[Spatial locality] on the other hand is used to refer to the concept that memory, which is spatially close to recently accessed memory, has a higher probability of being accessed itself.
\end{description}

\noindent
But why is all that useful? Well, it is due to the \em replacement policy \em of caches. As explained, caches will run full, and at that instance, they will have to evict a cache line. But how do they decide? Well, one common replacement policy is \em LRU (Least Recently Used)\em , which means that they choose the cache line whose access is the most far away in time. This is why temporal locality comes in handy, as it increases the chances that our data lies in cache, and thus reduces the risk of a cache miss.

@@

    Both matrices $a$ and $b$ are stored in row-order fashion. Hence, when running the left algorithm, $a$ has good spatial locality, whereas $b$'s is horrendous, as it leads to a cache miss on every access. The idea now is to find a tradeoff between these two. Let us analyze the access pattern of the two algorithms:
